Parallel Level-Set Methods on Adaptive Tree-Based Grids

### 2016-02-15 paper/peer-reviews/response_letter.tex

\begin{itemize} 
\item ``\textit{It took me a while to understand Algorithms 1 and 2. It is
  recom- mended that the authors’ pseudocode be closer to a language that many
people are familiar with like C++ or Fortran 90 (or perhaps, the authors’ are
using a new language that this referee is not aware of?).}''

Our library is coded in `\texttt{c++}' (with a mix of little `\texttt{c}' code).
We have made adjustments to the to algorithms to make them easier to understand.
However, we believe it is best not to directly use a `\texttt{c++}' syntax to
keep the pseudo-code language-agnostic and accessible to all readers.
\\
\item ``\textit{In Algorithm 2, what is the definition of “rank”? “owner’s
  rank”? “mpirank”? “st”?}''

There were multiple references to the ``\textit{rank}'' in the pseudo-code that 
would have caused confusion. We have made things clear by explicitly referring 
to ``\textit{owners\_rank}'' and ``\textit{mpirank}''. Also in the previous
pseudo-code ``\textit{st}'' was meant to indicate ``\texttt{MPI\_STATUS}''. Again
we have reorganized the pseudo-code to make this point more clear. We have also
added more comments and text to the caption to explain things in more details.
\\
\item ``\textit{In algorithm 2, if X is contained within the mesh on a given
  processor, does that mean all the points needed in order to derive the value
at X are also on the same processor? What if the points in the interpolatory
stencil of X belong to multiple processors? Which processor does the actual
interpolation? Is it the processor that needs the value at X? or is it the
processor that owns X?}''

Since the Z-curve defines non-overlapping partition of the grid in parallel, it
is guaranteed that the rank found by the search operation in line 3 of algorithm
2, actually owns the interpolation point, i.e. there exists a \textit{local
cell} \texttt{c} belonging to process \texttt{owners\_rank} which contains the
interpolation point. This is because the search operation uses a binary search
on the Z-curve to find the remote \texttt{owners\_rank}.

Our method for interpolation only requires that the values of the function (and
optionally its second derivatives if using multi-quadratic interpolation) be
defined on all vertices of cell \texttt{c}. This, in turn is guaranteed by the
fact that the boundary of each process is covered by a layer of ghost cells that
have updated the ghost values at all their vertices before performing the
interpolation. As a result, all the vertices of cell \texttt{c}, whether local
or ghost, have valid values that can be used for interpolation.
\\
\item ``\textit{In section 4.2, the authors should do a standard 3D reversible
  test when checking speed-up. A standard test allows researchers to compare
the new method to previous schemes. Also the authors should report the
numerical error (e.g. the symmetric difference error or the level set error in
proximity to the zero level set). For a reversible problem, the exact solution
is known at the end time. So in Table 2, an extra column should be added to
both the (a) and (b) parts which has the error (it is assumed that the error is
independent of p?). For the reversible problem, the zero level set isosurface
should be displayed at maximum deformation t = T/2 and at the end time t = T
for some representative values of “CFL” and lmax.}''

### 2016-02-15 paper/peer-reviews/response_letter.tex

\begin{document} 
\begin{letter} {Editorial Office, \\ Journal of Computational
Physics} 
\setkomavar{subject}{Re: JCOMP-D-15-00846}

\opening{Dear Editor,} 
We would like to appreciate the invaluable comments made by the reviewers. We
have made adjustments to our article, titled `\textbf{Parallel level-set methods
on adaptive tree-based grids}', to address the issues raised by the reviewers as
detailed below.

\begin{itemize} 
\item \textbf{Reviewer \#1} \\ 
The first reviewer has suggested to include ``\textit{some basic library codes
in appendix, each matching its level set algorithm}''. Indeed it is important
for the scientific community to share knowledge and facilitate ways for
reproducing scientific work for others. In the case of computational science,
unfortunately, this is a very daunting task that is not always possible.  Many
scientific codes are highly specialized and are often too technical to be fitted
into a paper.

Specifically, in our case, our library amounts to nearly 30,000 lines of code,
spread across more than 50 files and with complex dependencies. As such, it is
practically not viable to include sample codes in the paper. Moreover, we
believe inclusion of actual code, which involves many technical details, e.g.
error checking, profiling tools, comments, etc., is not scientifically
educational and may further alienate the reader from the main message of the paper.

Instead, and in accordance to the second reviewer who has raised related issues,
we have opted to revise the pseudo-codes in the algorithms to make them more
clear. We believe that the modifications should make the algorithms easier to
understand and more useful to the readers.

@@

\item ``\textit{Also in section 4.2, how frequently is reinitialization carried
  out? What is the pseudo time step $\Delta \tau$? What is $\tau_{stop}$? If
there is no reinitialization for this test, it is requested that the
semi-Lagrangian algorithm together with reinitialization be analyzed in this
section too}''

Previous studies have shown that the adaptive time-stepping method that is used
in this article greatly improves the convergence to steady state solution of
the reinitialization equation (c.f. ref 33 in main text). We have generally
observed that a few, typically 20, iterations of the algorithm is enough for
the solution to converge within a small band (e.g. one smallest cell in the
diagonal direction, i.e. $|\phi| < \sqrt{2}\Delta x_{min}$) around the
interface. As a result we solve the reinitialization equation not to a final
time (since each cell is using a different time-step) but to a fixed number of
iterations. We have added a new paragraph in section 3.3 that clarifies this
approach.

As for the test in section 4.2, there is no reinitialization involved. This is
because we are testing the effects of various parameters on the scalability of
a single iteration of the semi-Lagrangian. We believe, however, that the
reviewer is absolutely correct about the need to have a test of the entire
algorithm. However, we argue that this role is already fulfilled by the Stefan
test toward the end of the article.

The Stefan test of section 5 is a complete and non-trivial example and
represents a typical scenario that one should expect when using our algorithms
in practice. This example serves, not only as a demonstration of what our
algorithms can achieve, but also as a scalability test of all different
components, such as interpolation, semi-Lagrangian advection, reinitialization,
solution of implicit linear system, and extension of this solution over the
interface. It also compares these costs with costs associated with grid
operations in its entirety which includes generation, refinement and
coarsening, and partitioning of the forest. As figure 10, and table 7,
illustrate, both the semi-Lagrangian and the reinitialization scale well in a
real scenario. Thus we hope the results of this test is satisfactory enough for
the reviewer.
\end{itemize} 
\end{itemize}

### 2016-02-15 paper/tex/conclusion.tex
Small edits
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

In particular we showed that the scalability of the semi-Lagrangian algorithm, depends on the CFL number. Great scalability is observed for intermediate CFL numbers, e.g. $\text{CFL}\sim10$. At higher CFL numbers, however, the departure points are potentially further dispersed across processors, which limits the scalability. This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account. A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.\ by using a forward-in-time integration of grid points. Such an estimate could also reduce the number of semi-Lagrangian iterations. These ideas are postponed for further investigations. We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation. Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors.

Finally, an application of these algorithms is presented in modeling the solidification process by solving a Stefan problem. This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated practically using the domain decomposition techniques on uniform grids. We believe that our findings can serve as a basis to simulating a wide range of multi-scale problems and free body problems.

\section*{Acknowledgment} 
The funding for this project was provided by ONR N00014-11-1-0027.
We would also like to thank the technical staff at the Texas Advanced Computing
Center (TACC) and the developer community of PETSc library for their valuable
suggestions during this project.
The third author acknowledges support by the Hausdorff Center for Mathematics
(HCM) at Bonn University,
% and the Transregio 32 research collaborative, both
an initiative funded by the German Research Foundation (DFG).

>>>>

In particular we showed that the scalability of the semi-Lagrangian algorithm, depends on the CFL number. Great scalability is observed for intermediate CFL numbers, e.g. $\text{CFL}\sim10$. At higher CFL numbers, however, the departure points are potentially further dispersed across processors, which limits the scalability. This is because the domain decomposition technique used here is based on the Z-ordering of cells and does not take the velocity field information into account. A possible remedy for this problem could be assigning weights to cells based on some estimate of the grid structure after one step of the advection algorithm, e.g.\ by using a forward-in-time integration of grid points. Such an estimate could also reduce the number of semi-Lagrangian iterations. These ideas are postponed for further investigations. We have also presented a simple parallelization technique for the reinitialization algorithm based on the pseudo-time transient formulation. Both the semi-Lagrangian and the reinitialization algorithms show good scalability up to 4096 processors, the current limit of our account.

Finally, an application of these algorithms is presented in modeling the solidification process by solving a Stefan problem. This application clearly illustrates the applicability of our algorithms to complex multi-scale problems that cannot be treated practically using the domain decomposition techniques on uniform grids. We believe that our findings can serve as a basis to simulating a wide range of multi-scale and free boundary problems.

\section*{Acknowledgment} 
The funding for this project was provided by ONR N00014-11-1-0027. This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number
ACI-1053575. The authors acknowledge the Texas Advanced Computing
Center (TACC) at the University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper, URL: http://www.tacc.utexas.edu. We would also like to thank the developer community of PETSc library for their valuable
suggestions during this project.
The third author acknowledges support by the Hausdorff Center for Mathematics
(HCM) at Bonn University,
% and the Transregio 32 research collaborative, both
an initiative funded by the German Research Foundation (DFG).

### 2016-02-15 paper/tex/parallel_algorithms.tex
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

The advection of an irregular boundary using the semi-Lagrangian method is the perfect candidate to demonstrate the accuracy of our implementation as it makes use of the interpolation routine as well as the reinitialization procedure. We select the benchmark problem proposed in \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. Consider a sphere centered at $(0.35,0.35,0.35)$ and with radius $0.15$ in a domain $[0,1]^3$ and deformed under the divergence free velocity field
\be \label{eq:velo_enright}
\begin{cases}
u(x,y,z) = 2 \sin^2(\pi x)\sin(2 \pi y)\sin(2 \pi z), \\
v(x,y,z) = - \sin^2(\pi y)\sin(2 \pi x)\sin(2 \pi z), \\
w(x,y,z) = - \sin^2(\pi z)\sin(2 \pi x)\sin(2 \pi y).
\end{cases}
\ee
forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$. With set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest. We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases. The results are reported in table \ref{tab:accuracy} and figure \ref{fig:accuracy} shows a visualization of the sphere at the initial time, at $t_{1/2}$ when the deformation is maximal, and at the final time. The results are consistent with those reported previously in \cite{Min;Gibou:07:A-second-order-accur} and indicate second order accuracy for the mass loss.

@@

\begin{figure}[htbp]
\begin{center}
\begin{tabular}{ccccc}
\includegraphics[width=.28\textwidth]{figures/enright_0.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_1.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_2.png}
\end{tabular}
\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks. The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes. These results were obtained on the Comet supercomputer with 960 processes (40 compute nodes).} \label{fig:accuracy}
\end{center}
\end{figure}

>>>>

The advection of an irregular boundary using the semi-Lagrangian method is the perfect candidate to demonstrate the accuracy of our implementation as it makes use of the interpolation routine as well as the reinitialization procedure. We select the benchmark problem proposed in \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. Consider a sphere centered at $(0.35,0.35,0.35)$ and with radius $0.15$ in a domain $[0,1]^3$ and deformed under the divergence free velocity field
\be \label{eq:velo_enright}
\begin{cases}
u(x,y,z) = 2 \sin^2(\pi x)\sin(2 \pi y)\sin(2 \pi z), \\
v(x,y,z) = - \sin^2(\pi y)\sin(2 \pi x)\sin(2 \pi z), \\
w(x,y,z) = - \sin^2(\pi z)\sin(2 \pi x)\sin(2 \pi y).
\end{cases}
\ee
forward in time until $t_{1/2}=1$ and then backward to its original state at $t_f=2$. We set the time step to $\Delta t = 5 \Delta x_{min}$, where $\Delta x_{min}$ is the size of the smallest cell in the forest. The level-set function is reinitialized at every time step by applying 20 iterations of the reinitialization procedure. We monitor the volume loss and the error in the interface location at the final time, when the original shape should be recovered, as the finest resolution of the forest increases. The results are reported in table \ref{tab:accuracy} and figure \ref{fig:accuracy} shows a visualization of the sphere at the initial time, at $t_{1/2}$ when the deformation is maximal, and at the final time. The results are consistent with those reported previously in \cite{Min;Gibou:07:A-second-order-accur} and indicate second order accuracy for the mass loss.

@@

\begin{figure}[htbp]
\begin{center}
\begin{tabular}{ccccc}
\includegraphics[width=.28\textwidth]{figures/enright_blue_0.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_blue_1.png}
&
\hspace{.1cm}
&
\includegraphics[width=.28\textwidth]{figures/enright_blue_2.png}
\end{tabular}
\caption{Visualization of the deformation undergone by the sphere for the Enright's test at time 0, $t_{1/2}$ and $t_f$ and colored with the processes ranks. The forest's finest resolution is equivalent to a uniform grid with $2048^3$ cells and contains between 17 (at initial and final times) and 63 (at $t_{1/2}$) million nodes. These results were obtained on the Comet supercomputer with 960 processes (40 compute nodes).} \label{fig:accuracy}
\end{center}
\end{figure}

### 2016-02-15 paper/tex/scaling_results.tex
update enright figure with blue color, add number of iterations for reinitilization, add citations for XSEDE, stampede and comet

%!TEX root = draft.tex
\section{Scaling results} \label{sec:scaling}
In this section we present some results that demonstrate the scalability of our algorithms. All of our tests were ran on the Stampede cluster at the Texas Advanced Computing Center (TACC) where we are limited to $4096$ cores at most. Each node of Stampede has 2 eight-core Xenon E5-2680 processes clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card. Unless mentioned otherwise, in all the tests we have used all 16 cores of every node. Finally, in all cases we report the maximum wall time recorded using PETSc's logging interface which has a temporal resolution of roughly $0.1 \: \mu s$.

>>>>

%!TEX root = draft.tex
\section{Scaling results} \label{sec:scaling}
In this section we present some results that demonstrate the scalability of our algorithms. All of our tests were ran on the Stampede cluster at the Texas Advanced Computing Center (TACC), a resource accessible through the Extreme Science and Engineering Discovery Environment (XSEDE) \cite{TownsCockerillDahanEtAl:14:XSEDE:AcceleratingScientific}, where we are limited to $4096$ cores at most. Each node of Stampede has 2 eight-core Xenon E5-2680 processes clocked at 2.7 GHz with 32 GB of DDR3-1600 MHz memory and interconnected using an InfiniBand network card. Unless mentioned otherwise, in all the tests we have used all 16 cores of every node. Finally, in all cases we report the maximum wall time recorded using PETSc's logging interface which has a temporal resolution of roughly $0.1 \: \mu s$.
